---
title: "ACM Final Project"
author: "Alyster Alcudia, Cynthia Gan, Mihai Sirbu"
date: "5/2/2015"
output: pdf_document
---

```{r, echo=FALSE, include=FALSE}
for(i in c("corrplot", "car", "dplyr", "tidyr", "ggplot2", "gridExtra", "agricolae")){
    if( ! require( i , character.only = TRUE ) ){
      #  If package was not able to be loaded then re-install
      install.packages( i , dependencies = TRUE )
      library( i , character.only = TRUE )
    }
}

```
Data
====

We are looking at data about students in secondary schools in Portugal. It contains information about student grades, demographics, and other social factors, collected through school reports and questionnaires.  

We specifically look at grades for Portuguese language class. School years in this set are split into three periods, and as a result there are three grades. The grade issued after the third period is the final grade. 

After cleaning the data (below) by renaming columns and making responses more meaningful our data consists of 649 observations and 33 variables. The end variables are as follows: 

###Variables in Table 1: 

* school - student's school ("Gabriel Pereira" or "Mousinho da Silveira")
* sex - student's sex ("F" - female or "M" - male)
* age - student's age (from 15 to 22)
* address - student's home address type ("Urban" or "Rural")
* fam_size - family size ("<=3" - less than or equal to three or ">3" - greater than 3)
* parent_status - parent's cohabitation status ("together" or "apart")
* mom_edu - mother's education (0 - none,  1 - primary education (4th grade), 2 – 5th to 9th grade, 3 – secondary education or 4 – higher education)
* dad_edu - father's education (0 - none,  1 - primary education (4th grade), 2 – 5th to 9th grade, 3 – secondary education or 4 – higher education)
* mom\_job - mother's job ("teacher", "health" care related, civil "services" (e.g. administrative or police), "at_home" or "other")
* dad\_job - father's job ("teacher", "health" care related, civil "services" (e.g. administrative or police), "at_home" or "other")
* reason - reason to choose this school (close to "home", school "reputation", "course" preference or "other")
* guardian - student's guardian ("mother", "father" or "other")
* traveltime - home to school travel time (1 - <15 min., 2 - 15 to 30 min., 3 - 30 min. to 1 hour, or 4 - >1 hour)
* studytime - weekly study time (1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours)
* failures - number of past class failures (1, 2, 3, or 4 (with 4 representing 4 or more))
* extra\_school\_support - extra educational support (TRUE or FALSE)
* fam\_edu\_sup - family educational support (TRUE or FALSE)
* paid\_extra\_classes - extra paid classes within the course subject (Portuguese) (TRUE or FALSE)
* extracurricular_activities - extra-curricular activities (TRUE or FALSE)
* nursery - attended nursery school (TRUE or FALSE)
* higher - wants to take higher education (TRUE or FALSE)
* internet - Internet access at home (TRUE or FALSE)
* romantic - with a romantic relationship (TRUE or FALSE)
* family_relationship - quality of family relationships (from 1 - very bad to 5 - excellent)
* freetime - free time after school (from 1 - very low to 5 - very high)
* going\_out\_amount - amount going out with friends (from 1 - very low to 5 - very high)
* weekday\_alcohol\_cons - workday alcohol consumption (from 1 - very low to 5 - very high)
* weekend\_alcohol\_cons - weekend alcohol consumption (from 1 - very low to 5 - very high)
* health - current health status (from 1 - very bad to 5 - very good)
* absences - number of school absences (from 0 to 93)
* __grade_one__ - first period Portuguese grade (from 0 to 20)
* __grade_two__ - second period Portuguese grade (from 0 to 20)
* __final_grade__ - final Portuguese grade (from 0 to 20)

###Next Steps

We plan on using linear regression to infer what factors influence student achievement, as measured by their final grades. We will produce an R-Markdown file with our findings as an end result.


Data Tidying
============

###Read In Data

```{r}
data = as.tbl(read.csv("student-por.csv", sep=";", stringsAsFactors=TRUE))
```

###Rename the Variables

```{r}

rename_cols <- c(4:12,16:21,23:24,26:28,31:33)

colnames(data)[rename_cols] = c("urban_rural",
                                "fam_size",
                                "parents_cohabit",
                                "mom_edu",
                                "dad_edu",
                                "mom_job",
                                "dad_job",
                                "school_reason",
                                "student_guardian",
                                "extra_school_support",
                                "fam_edu_support",
                                "paid_extra_classes",
                                "extracurricular_activities",
                                "attended_nursery",
                                "wants_higher_education",
                                "has_romantic_partner",
                                "family_relationship",
                                "going_out_amount",
                                "weekday_alcohol_cons",
                                "weekend_alcohol_cons",
                                "grade_one",
                                "grade_two",
                                "final_grade")

```

###Meaningful Variable Values

```{r}
data <-  data %>% 
  mutate(school = ifelse(school=='GP', 'Gabriel Pereira', 'Mousinho da Silveira')) %>%
  mutate(urban_rural = ifelse(urban_rural=='R', 'Rural', 'Urban')) %>% 
  mutate(fam_size = ifelse(fam_size=='GT3', '>3', '<=3')) %>%
  mutate(parents_cohabit = ifelse(parents_cohabit=='A', 'Apart', 'Together')) %>%
  mutate_each_(funs(ifelse(.=='yes', TRUE, FALSE)), 
               c("extra_school_support", "fam_edu_support", "paid_extra_classes", 
                 "extracurricular_activities", "attended_nursery", 
                 "wants_higher_education", "internet", "has_romantic_partner"))
  
```

Data Exploration
================

Basic view of the data
----------------------
```{r}
dim(data)
```

The tidy data is in 1 table, with 649 observations of 33 variables. 



```{r}
summary(data)
```

We have data from two schools, mostly from Gabriel Pereira. The sample is largely urban, with parents living together and a family size greater than 3. Most parents don't tend to have completed secondary education. A large majority, `r mean(data$wants_higher_education)`, wants higher education despite that. This is reflected in that the mean number of absences if `r mean(data$absences)` out of a possible 93. 

Correlation  
-----------

Correlation betweens between different grades and
numeric variables
```{r}
data %>%
  select(one_of("age","failures", "family_relationship",
                "freetime","going_out_amount",
                "weekday_alcohol_cons",
                "weekend_alcohol_cons","absences",
                "grade_one","grade_two",
                "final_grade")) %>%
  cor() %>%
  corrplot(method="circle",type="lower")

```

In most cases, the correlations are not particularly strong. However there is medium and negative correlations between failures and grades. This makes intuitive sense given that a person who has failed previously probably has poorer school performance. Alcohol consumption during the week has a strong positive correlation, which is not surprising. Finally, the three types of grades are very strongly correlated, with a correlation of anywhere between 0.83 to 0.92. This strong correlation suggests that we may be able to focus on one of the three types of grades (e.g. final grade) when trying to understand the factors that influence student performance. 


What about correlations between family/home attributes? Any strong correlations between such variables can lead to misleading estimates when modeling the data. In this analysis, we will consider the `fam_size`, `parents_cohabit` and `fam_edu_support` variables with respect to `family_relationship`.

```{r}
tbl = table(data$family_relationship, data$fam_size)
tbl

chisq.test(tbl)
```

```{r}
tbl = table(data$family_relationship, data$parents_cohabit)
chisq.test(tbl)
```

```{r}
tbl = table(data$family_relationship, data$fam_edu_support)
chisq.test(tbl)
```

The above chi-squared tests result in high p-values; there is not enough evidence to reject the null hypothesis. Therefore, these variables are independent.

What about the job types and education levels for each parent?

```{r}
edutbl = table(data$mom_edu, data$dad_edu)
chisq.test(edutbl)

jobtbl = table(data$mom_job, data$dad_job)
chisq.test(jobtbl)
```

The p-value is small enough to reject the null hypothesis. That is, the job types and education levels for each parent are highly correlated. Although there is a warning from `chisq.test` (most likely because of the small cell values of `tbl`), clear trends can be seen looking at the values of `tbl`.

Since `mom_edu` and `dad_edu` are numerical values, we can compute the correlation between them.

```{r}
cor(data$mom_edu, data$dad_edu)
```

The above code shows that on average, `mom_edu` and `dad_edu` are positively correlated. Therefore, including both variables in a data model can lead to misleading estimates.

Gender and Grade  
----------------


```{r}
table(data$sex)
```

There number of male and female students are approximately equal, which make them good groups to compare. 

```{r}
long_grades <- data %>% 
                gather(Grade_Type, Grade, 
                       grade_one:final_grade)
  
long_grades %>%
  ggplot(aes(x=Grade, fill=sex))+
  geom_density(alpha=0.2, binwidth = 2, position = 'identity')+
  facet_wrap(~Grade_Type)+
  ggtitle("Density Plot across the three grades as a function of gender")
```

The plots don't indicate obvious differenes between genders in terms of grade performance. Let's just do a t-test between genders to see if we're missing something.

```{r}
males_one <- long_grades %>% filter(sex=="M",Grade_Type=="grade_one")
females_one <- long_grades %>% filter(sex=="F",Grade_Type=="grade_one")

t.test(x=males_one$Grade,y=females_one$Grade,alternative="two.sided")
```
The mean grade of female students in period one is signficantly higher than the mean grade of male students. The t-test gives a p-value 0.007, which means that it is extremely unlikely that the difference of the means of the two groups is actually 0.

``` {r}
males_two <- long_grades %>% filter(sex=="M",Grade_Type=="grade_two")
females_two <- long_grades %>% filter(sex=="F",Grade_Type=="grade_two")

t.test(x=males_two$Grade,y=females_two$Grade,alternative="two.sided")
```

The mean grade of female students in period two is also signficantly higher than the mean grade of male students, at `r mean(females_two$Grade)` compared to `r mean(males_two$Grade)`. The probability that the difference of the means of the two groups is actually 0 is about 0.007, which means that it is extremely unlikely that we are seeing noise.


```{r}
males_final <- long_grades %>% filter(sex=="M",Grade_Type=="final_grade")
females_final <- long_grades %>% filter(sex=="F",Grade_Type=="final_grade")

t.test(x=males_final$Grade,y=females_final$Grade,alternative="two.sided")

```


The mean final grade of female students is `r mean(females_two$Grade)`, and the mean final grade of male is `r mean(males_two$Grade)`. Since the p-value is extremely low, at about 0.001, the difference of `r mean(females_two$Grade) - mean(males_two$Grade)` is a statistiscally significant difference.

###Conclusion of Gender and Grades

As it turns out, in all three cases, females had a signficantly higher grade than males as all three p values were less than 0.05.  


Let's see if there are any other gender differences for some of the other variables! (Note: in cases where the scale is from 1 to 5, 1 represents lowest (e.g. very amount of studying) whereas 5 represents highest (e.g. very high amount of studying))


Misc Variable Relations
-----------------------
```{r, out.width='0.75\\maxwidth', out.height='0.45\\maxwidth'}
data %>% group_by(sex, has_romantic_partner) %>% tally() %>% mutate(freq = n/sum(n)) %>%
  ggplot(aes(x=sex,y=freq, fill=has_romantic_partner))+
  geom_bar(stat="identity",position="dodge")+
  ggtitle("Presence of a Romantic Partner\nas a function of Sex (Rel. Freq)")+
  ylim(0,1)
```

_Females seem to be more likely to have romantic partners than males._

\pagebreak

```{r, out.width='0.75\\maxwidth', out.height='0.45\\maxwidth'}
data %>% group_by(sex, urban_rural) %>% tally() %>% mutate(freq = n/sum(n)) %>%
  ggplot(aes(x=sex,y=freq, fill=urban_rural))+
  geom_bar(stat="identity",position="dodge")+
  ggtitle("Distribution of Rural/Urban location as a function of sex (Rel. Freq)")+
  ylim(0,1)

```

_Gender and Urban/Rural appear to be independent of each other._

```{r, out.width='0.75\\maxwidth', out.height='0.45\\maxwidth'}

data %>% group_by(sex, fam_size) %>% tally() %>% mutate(freq = n/sum(n)) %>%
  ggplot(aes(x=sex,y=freq, fill=fam_size))+
  geom_bar(stat="identity",position="dodge")+
  ggtitle("Student Family Size as a function of sex (Rel. Freq)")+
  ylim(0,1)

```

_Gender doesn't appear to have a significant affect on family size._

```{r, out.width='0.75\\maxwidth', out.height='0.45\\maxwidth'}
data %>% group_by(sex, study=factor(studytime)) %>% tally() %>% mutate(freq = n/sum(n)) %>%
  ggplot(aes(x=sex,y=freq, fill=study))+
  geom_bar(stat="identity",position="dodge")+
  ggtitle("Student Study Time as a function of Sex (Rel. Freq)")+
  ylim(0,1)
```
_Females seem to study more than males._

```{r, out.width='0.75\\maxwidth', out.height='0.45\\maxwidth'}
data %>% group_by(sex, freetime=factor(freetime)) %>% tally() %>% mutate(freq = n/sum(n)) %>%
  ggplot(aes(x=sex,y=freq, fill=freetime))+
  geom_bar(stat="identity",position="dodge")+
  ggtitle("Student Free Time as a function of Sex (Rel. Freq)")+
  ylim(0,1)
```

_Males seem to have more free time than Females._


```{r, out.width='0.75\\maxwidth', out.height='0.45\\maxwidth'}
data %>% group_by(sex, weekend_alcohol_consumption=factor(weekend_alcohol_cons)) %>% tally() %>% mutate(freq = n/sum(n)) %>%
  ggplot(aes(x=sex,y=freq, fill=weekend_alcohol_consumption))+
  geom_bar(stat="identity",position="dodge")+
  ggtitle("Weekend Alcohol Consumption\nas a function of Sex (Rel. Freq)")+
  ylim(0,1)
```

_Males tend to drink more alcohol on the weekend than females._

###Misc Variable Relationship Conclusions   

Looking at the different plots, visual inspection shows a difference (unclear if significant) in relative frequencies on the basis of sex in the following variables: weekend alcohol consumption (with men drinking more), presence of romantic partners (with females more likely having a partner), and amount of study and free time (with females seeming to study more and males having more free time). Urban/Rural divisions and family size seemed to have less differences (again by visual inspection) than the other graphs. 

Final Grades and Gender
-----------------------

We've looked at the relative frequencies using barplots. What if we examined relationships using scatterplots grouping by sex? 

```{r}
fg_fg_abs_sex = data %>%
  ggplot(aes(x = absences, y = final_grade, color=sex)) + geom_point() + geom_smooth(method=lm)+
  ggtitle("Final Grade as a\nfunction of absences\ngrouped by sex")+
  theme(aspect.ratio = 1)

fg_fg_std_sex = data %>%
  ggplot(aes(x=studytime, y = final_grade, color=sex)) + geom_point() +geom_smooth(method=lm)+
  ggtitle("Final Grade as a\nfunction of study time\ngrouped by sex")+
  theme(aspect.ratio = 1)

grid.arrange(fg_fg_abs_sex, fg_fg_std_sex, nrow=1)
```

There doesn't appear to be any clear noticeable differences in the slopes of the lines for either plot, suggesting that if we used study time or absences to predict/understand final grade, these variables wouldn't have an interaction with sex.



Parental Education Levels
-------------------------

```{r}
mom_edu_finals <- data %>%    
    mutate(mom_edu_fac = factor(mom_edu, levels=seq(0,4), labels=c("None",
      "4th grade", "5th-9th\ngrade", "Secondary\neducation", 
      "Higher\neducation"), ordered=TRUE)) %>%
    group_by(mom_edu_fac) %>%
    summarise(mean_est = t.test(final_grade)$estimate, 
              left_CI = t.test(final_grade)$conf.int[1], 
              right_CI = t.test(final_grade)$conf.int[2])

mom_edu_finals_plot <- ggplot(mom_edu_finals, aes(x=mom_edu_fac)) +
    geom_point(aes(y=mean_est), size=3, color='darkgreen') +
    geom_errorbar(aes(ymin=left_CI, ymax=right_CI), width=0.1) +
    labs(title="Estimated Mean Final Grade\nand 95% Confidence Interval\n
        by Mother's Education Level", x="Mother's Education Level", 
        y="Estimated Mean Final Grade")+
    ylim(5,15) + theme(axis.text.x=element_text(angle=50, vjust=0.5))

dad_edu_finals = data %>%
    mutate(dad_edu_fac = factor(dad_edu, levels=seq(0,4), 
                                labels=c("None", "4th grade", "5th-9th\ngrade", 
                                         "Secondary\neducation", "Higher\neducation"), 
                                ordered=TRUE)) %>%
    group_by(dad_edu_fac) %>%
    summarise(mean_est = t.test(final_grade)$estimate, 
              left_CI = t.test(final_grade)$conf.int[1], 
              right_CI = t.test(final_grade)$conf.int[2])

dad_edu_finals_plot <- ggplot(dad_edu_finals, aes(x=dad_edu_fac)) +
    geom_point(aes(y=mean_est), size=3, color='darkgreen') +
    geom_errorbar(aes(ymin=left_CI, ymax=right_CI), width=0.1) +
    labs(title="Estimated Mean Final Grade\nand 95% Confidence Interval\n
        by Mother's Education Level", x="Mother's Education Level", 
        y="Estimated Mean Final Grade")+
    ylim(5,15) + theme(axis.text.x=element_text(angle=50, vjust=0.5))


grid.arrange(mom_edu_finals_plot, dad_edu_finals_plot, nrow=1)
```

_If you discount parents with no education, there is a clear relationship between parental education and child academic achievement_

###Mother's Education  

Suprisingly, the estimated mean final school for students whose mother has no education is almost as high as students whose mother has education up to secondary school. However, The confidence interval for this estimate is extremely high compared to the other estimates, spanning `r mom_edu_finals$right_CI[1] - mom_edu_finals$left_CI[1]` compared to the others, which tend to span about 1 point (`r mom_edu_finals$right_CI[2:5]-mom_edu_finals$left_CI[2:5]` respectively). This is likely because there are very few mothers with no education, which makes the mean very susceptible to outliers. One student who does unusually well could highly influence the estimated mean.  

Aside from the estimate for student's with mother's with no education, estimates for the mean final grades steadily increase as the amount of education their mother recieved increases. However, the difference between the mean final grade for a student with a mother with higher education and a student with a mother with 4th grade education is only `r mom_edu_finals$mean_est[5] - mom_edu_finals$mean_est[1]`, which is not very large.

```{r}
HSD.test(aov(final_grade ~ mom_edu, data=data), "mom_edu", group=TRUE, 
         alpha = 0.05, console=TRUE)
```
The Tukey HSD test on final grades by mother's education level finds that students whose mother has had higher education fare significantly better than students who don't. Mothers with a secondary education are also significantly better than mothers with only 4th grade education, but no other comparisons have significant differences.


###Father's Education  

The estimated mean final grades for each level of father education shows very similar trends as the mean final grades for levels of mother education. However, the effects appear to be slightly stronger. The difference between the higher education and 4th grade in fathers is `r dad_edu_finals$mean_est[5] - dad_edu_finals$mean_est[1]`, compared to `r mom_edu_finals$mean_est[5] - mom_edu_finals$mean_est[1]` in mothers.  

```{r}
HSD.test(aov(final_grade ~ dad_edu, data=data), "dad_edu", group=TRUE, 
         alpha = 0.05, console=TRUE)
```

The Tukey HSD Test on final grades by father's education level was much less conclusive than that of the mother's. The effect of the different education levels appears to be more incremental. Having a father with a higher education is associated with a significantly better score than having one who did not complete secondary education. Having a father with second education is associated with significantly better final scores than havign one with 4th grade education. 

###Overall  


These exploratory plots suggest there may be a relationship between parental education levels and academic achievement. There are many possible underlying explanations for this possible relation, such as a stronger parental emphasis on education, or being able to provide higher quality support. 


Parental Employment 
-------------------

```{r}
mom_job_finals <- data %>% 
  group_by(mom_job) %>% 
  summarise(mean_est = t.test(final_grade)$estimate, 
            left_CI = t.test(final_grade)$conf.int[1], 
            right_CI = t.test(final_grade)$conf.int[2]) 

mom_job_finals_plot = ggplot(mom_job_finals, aes(x=mom_job)) + 
  geom_point(aes(y=mean_est), size=3, color='darkgreen') + 
  geom_errorbar(aes(ymin=left_CI, ymax=right_CI), width=0.1) + 
  labs(title="Estimated Mean Final Grade\nand 95% Confidence Interval\n
       by Mother's Job", x="Mother's Job", y="Estimated Mean Final Grade")+ 
  ylim(5,15) + 
  theme(axis.text.x=element_text(angle=50, vjust=0.5))

dad_job_finals <- data %>% 
  group_by(dad_job) %>% 
  summarise(mean_est = t.test(final_grade)$estimate, 
            left_CI = t.test(final_grade)$conf.int[1], 
            right_CI = t.test(final_grade)$conf.int[2]) 

dad_job_finals_plot = ggplot(dad_job_finals, aes(x=dad_job)) + 
  geom_point(aes(y=mean_est), size=3, color='darkgreen') + 
  geom_errorbar(aes(ymin=left_CI, ymax=right_CI), width=0.1) + 
  labs(title="Estimated Mean Final Grade\nand 95% Confidence Interval\n
       by Father's Job", x="Father's Job", y="Estimated Mean Final Grade")+ 
  ylim(5,15) + 
  theme(axis.text.x=element_text(angle=50, vjust=0.5))

grid.arrange(mom_job_finals_plot, dad_job_finals_plot, nrow=1)
```

_Estimated final grades show similar trends across mother and father employment._

Health and teaching industries are associated with the highest education levels, which intuitively makes sense. They are both industries that require extensive education to qualify for, and previous plots suggested that there may be a relationship between parental education and academic performance.
Meanwhile, parents that stay at home correspond to worse student achievement. However, it is unclear whether  "at home" is purposeful or not. If "at home" includes parents who are unemployed but seeking work, there may be underlying factors for poorer academic achievement, such as a stressful home environment. 

###Mother's Employment

```{r}
HSD.test(aov(final_grade ~ mom_job, data=data), "mom_job", group=TRUE, 
         alpha=0.05, console=TRUE)
```
Based on the Tukey HSD test, we see that there are significant differences in final grade between the types of industries the student's mother works in. Mothers who work in teaching or health have students which significantly higher score than students who work in other or stay at home. However, it is not a clean split. Mothers who work in service have student's whose final grades are not significantly different from those in health, teaching or other. 

###Father's Employment

```{r}
HSD.test(aov(final_grade ~ dad_job, data=data), "dad_job", group=TRUE, 
         alpha=0.05, console=TRUE)
```

The split in the Tukey HSD test on father profession is slightly cleaner than that on mother profession. Teacher fathers correspond to significantly higher final scores than other, service, or at home fathers. Fathers in the latter three industries do not have statistically significant differences from each other, but are different from teacher fathers. However, fathers in the health profession are not significantly different from either group. 


###Overall

There is statistically siginficance to the type of employment of parents. Students with parents that are teachers, or mothers in health tend to have higher final grades than ohter students. 
