---
title: "ACM_Final_Proj"
author: "Alyster Alcudia, Cynthia Gan, Mihai Sirbu"
date: "5/2/2015"
output: pdf_document
---

```{r, echo=FALSE, include=FALSE}
install.packages("corrplot")
install.packages("dplyr")
install.packages("tidyr")
install.packages("ggplot2")

library(corrplot)
library(dplyr)
library(tidyr)
library(ggplot2)
```

```{r}
data = as.tbl(read.csv("student-por.csv", sep=";", stringsAsFactors = TRUE))
```

Let's rename the variables better!
```{r}

rename_cols <- c(4, 6, 11:12,16:21,23:24,26:28,31:33)

colnames(data)[rename_cols] = c("urban_rural",
                                "parents_cohabit",
                                "school_reason",
                                "student_guardian",
                                "extra_school_support",
                                "fam_edu_support",
                                "paid_extra_classes",
                                "extracurricular activites",
                                "attended_nursery",
                                "wants_higher_education",
                                "has_romantic_partner",
                                "family_relationship",
                                "going_out_amount",
                                "weekday_alcohol_cons",
                                "weekend_alcohol_cons",
                                "first_per_grade",
                                "second_per_grade",
                                "final_grade")

```
Basic view of the data
```{r}
dim(data)
summary(data)
```

Correlation betweens between different grades and
numeric variables
```{r}
data %>%
  select(one_of("age","failures", "family_relationship",
                "freetime","going_out_amount",
                "weekday_alcohol_cons",
                "weekend_alcohol_cons","absences",
                "first_per_grade","second_per_grade",
                "final_grade")) %>%
  cor() %>%
  corrplot(method="number",type="upper")

```

In most cases, the correlations are not particularly strong. However there is medium and negative correlations between failures and grades. This makes intuitive sense given that a person who has failed previously probably has poorer school performance. Alcohol consumption during the week and during the week has a strong positive correlation, which is not surprising. Finally, the three types of grades are very strongly correlated with a correlation of anywhere between 0.83 to 0.92. This strong correlation suggests that we may be able to focus on one of the three types of grades (e.g. final grade) when trying to understand the factors that influence student performance. 

Let's examine if gender seems to play a role in terms of the grade
```{r}

table(data$sex)

long_grades <- data %>% 
                gather(Grade_Type, Grade, 
                       first_per_grade:final_grade)
  
long_grades %>%
  ggplot(aes(x=Grade, fill=sex))+
  geom_density(alpha=0.2, binwidth = 2, position = 'identity')+
  facet_wrap(~Grade_Type)+
  ggtitle("Density Plot across the three grades as a function of gender")
```

The plots don't indicate obvious differenes between genders in terms of grade performance. Let's just do a t.test between genders to see if we're missing something

```{r}
males_one <- long_grades %>% filter(sex=="M",Grade_Type=="first_per_grade")
females_one <- long_grades %>% filter(sex=="F",Grade_Type=="first_per_grade")

t.test(x=males_one$Grade,y=females_one$Grade,alternative="two.sided")

males_two <- long_grades %>% filter(sex=="M",Grade_Type=="second_per_grade")
females_two <- long_grades %>% filter(sex=="F",Grade_Type=="second_per_grade")

t.test(x=males_two$Grade,y=females_two$Grade,alternative="two.sided")

males_final <- long_grades %>% filter(sex=="M",Grade_Type=="final_grade")
females_final <- long_grades %>% filter(sex=="F",Grade_Type=="final_grade")

t.test(x=males_final$Grade,y=females_final$Grade,alternative="two.sided")

```
As it turns out, in all three cases, females had a signficantly higher grade than males as all three p values were less than 0.05. 

Let's see if there are any other gender differences for some of the other variables!

```{r}
data %>% group_by(sex, has_romantic_partner) %>% tally() %>% mutate(freq = n/sum(n)) %>%
  ggplot(aes(x=sex,y=freq, fill=has_romantic_partner))+
  geom_bar(stat="identity",position="dodge")+
  ggtitle("Presece of a Romantic Partner as a function of Sex (Rel. Freq)")

data %>% group_by(sex, urban_rural) %>% tally() %>% mutate(freq = n/sum(n)) %>%
  ggplot(aes(x=sex,y=freq, fill=urban_rural))+
  geom_bar(stat="identity",position="dodge")+
  ggtitle("Distribution of Rural/Urban location as a function of sex (Rel. Freq)")

data %>% group_by(sex, famsize) %>% tally() %>% mutate(freq = n/sum(n)) %>%
  ggplot(aes(x=sex,y=freq, fill=famsize))+
  geom_bar(stat="identity",position="dodge")+
  ggtitle("Student Family Size as a function of sex (Rel. Freq)")

data %>% group_by(sex, study=factor(studytime)) %>% tally() %>% mutate(freq = n/sum(n)) %>%
  ggplot(aes(x=sex,y=freq, fill=study))+
  geom_bar(stat="identity",position="dodge")+
  ggtitle("Student Study Time as a function of Sex (Rel. Freq)")

data %>% group_by(sex, freetime=factor(freetime)) %>% tally() %>% mutate(freq = n/sum(n)) %>%
  ggplot(aes(x=sex,y=freq, fill=freetime))+
  geom_bar(stat="identity",position="dodge")+
  ggtitle("Student Free Time as a function of Sex (Rel. Freq)")

data %>% group_by(sex, weekend_alcohol_consumption=factor(weekend_alcohol_cons)) %>% tally() %>% mutate(freq = n/sum(n)) %>%
  ggplot(aes(x=sex,y=freq, fill=weekend_alcohol_consumption))+
  geom_bar(stat="identity",position="dodge")+
  ggtitle("Weekend alcohol Consumption as a function of Sex (Rel. Freq)")
```

Looking at the different plots, visual inspection shows a difference (unclear if significant) in relative frequencies on the basis of sex in the following variables: weekend alcohol consumption (with men drinking more), presence of romantic partners (with females more likely having a partner), and amount of study and free time (with females seeming to study more and males having more free time). Urban/Rural divisions and family size seemed to have less differences (again by visual inspection) than the other graphs. 

We've looked at the relative frequencies using barplots. What if we examined relationships using scatterplots grouping by sex? 

```{r}
data %>%
  ggplot(aes(x = absences, y = final_grade, color=sex)) + geom_point() + geom_smooth(method=lm)+
  ggtitle("Final Grade as a function of absences grouped by sex")

data %>%
  ggplot(aes(x=studytime, y = final_grade, color=sex)) + geom_point() +geom_smooth(method=lm)+
  ggtitle("Final Grade as a function of study time grouped by sex")
```

There doesn't appear to be any clear noticeable differences in the slopes of the lines for either plot, suggesting that if we used study time or absences to predict/understand final grade, these variables wouldn't have an interaction with sex.
