---
title: "Supplmental_Results"
author: "Mihai Sirbu"
date: "Saturday, May 16, 2015"
output: html_document
---

I'm including this sheet because we may be doing a lot of random
analyses and only a subset of them make it into the final report. 

```{r, echo=FALSE, include=FALSE}
for(i in c("dplyr", "MASS", "ggplot2")){
    if( ! require( i , character.only = TRUE ) ){
      #  If package was not able to be loaded then re-install
      install.packages( i , dependencies = TRUE )
      library( i , character.only = TRUE )
    }
}
```

###Read in Data

```{r}
data = as.tbl(read.csv("student-por.csv", sep=";", stringsAsFactors=TRUE))
```

###Rename the Variables

```{r}

rename_cols <- c(4:12,16:21,23:24,26:28,31:33)

colnames(data)[rename_cols] = c("urban_rural",
                                "fam_size",
                                "parents_cohabit",
                                "mom_edu",
                                "dad_edu",
                                "mom_job",
                                "dad_job",
                                "school_reason",
                                "student_guardian",
                                "extra_school_support",
                                "fam_edu_support",
                                "paid_extra_classes",
                                "extracurricular_activities",
                                "attended_nursery",
                                "wants_higher_education",
                                "has_romantic_partner",
                                "family_relationship",
                                "going_out_amount",
                                "weekday_alcohol_cons",
                                "weekend_alcohol_cons",
                                "grade_one",
                                "grade_two",
                                "final_grade")
```

###Meaningful Variable Values

```{r}
data <-  data %>% 
  mutate(school = ifelse(school=='GP', 'Gabriel Pereira', 'Mousinho da Silveira')) %>%
  mutate(urban_rural = ifelse(urban_rural=='R', 'Rural', 'Urban')) %>% 
  mutate(fam_size = ifelse(fam_size=='GT3', '>3', '<=3')) %>%
  mutate(parents_cohabit = ifelse(parents_cohabit=='A', 'Apart', 'Together')) %>%
  mutate_each_(funs(ifelse(.=='yes', TRUE, FALSE)), 
               c("extra_school_support", "fam_edu_support", "paid_extra_classes", 
                 "extracurricular_activities", "attended_nursery", 
                 "wants_higher_education", "internet", "has_romantic_partner"))
  
```

```{r, results="hide"}
fit <- lm(final_grade~.-grade_one-grade_two,data=data)
step2 <- stepAIC(fit, direction="both")


summary(step2)
```

Someone should take a few momements to interpret these results...


The banding pattern of the residuals is WEIRD. We need to investigate why
this is. 
```{r}
fitted_res <- as.tbl(data.frame(fitted = fitted(step2), res = residuals(step2)))

data_fitted_res <- as.tbl(cbind(data, fitted_res))

#Normal
data_fitted_res %>%
  ggplot(aes(x=fitted, y=res)) +geom_point(alpha=0.5)

data_fitted_res %>%
  ggplot(aes(x=fitted, y=res, color=extra_school_support)) +geom_point(alpha=0.5)

```


```{r}
fit2 = lm(formula = final_grade ~ school * mom_edu * failures * extra_school_support * 
            wants_higher_education * weekday_alcohol_cons, data = data)
fit2b = lm(formula = final_grade ~ school * mom_edu * weekday_alcohol_cons * 
           (extra_school_support + failures), data = data)

fit2c = lm(formula = final_grade ~ school + mom_edu + failures + extra_school_support + 
             wants_higher_education + weekday_alcohol_cons + school:extra_school_support + 
             school:mom_edu:extra_school_support + school:weekday_alcohol_cons:extra_school_support + 
             school:mom_edu:weekday_alcohol_cons:extra_school_support, 
           data = data)
#pulled from siginficant terms of fit2b

fit2d = lm(final_grade ~ failures + wants_higher_education + school + extra_school_support + mom_edu + weekday_alcohol_cons + (school:extra_school_support)*(mom_edu*weekday_alcohol_cons), data=data)
#Factored from fit2c
#0.3288/0.3429 = 0.96% 








a1 = lm(final_grade ~ school, data=data)
a2 = lm(final_grade ~ mom_edu, data=data)
a3 = lm(final_grade ~ failures, data=data)
a4 = lm(final_grade ~ extra_school_support, data=data)
a5 = lm(final_grade ~ wants_higher_education, data=data)
a6 = lm(final_grade ~ weekday_alcohol_cons, data=data)
a7 = lm(final_grade ~ school*mom_edu, data=data)
a8 = lm(final_grade ~ school*failures, data=data)
a9 = lm(final_grade ~ school*extra_school_support, data=data)
a10 = lm(final_grade ~ school*wants_higher_education, data=data)
a11 = lm(final_grade ~ school*weekday_alcohol_cons, data=data)
a12 = lm(final_grade ~ mom_edu*failures, data=data)
a13 = lm(final_grade ~ mom_edu*extra_school_support, data=data)
a14 = lm(final_grade ~ mom_edu*wants_higher_education, data=data)
a15 = lm(final_grade ~ mom_edu*weekday_alcohol_cons, data=data)
a16 = lm(final_grade ~ failures*extra_school_support, data=data)
a17 = lm(final_grade ~ failures*wants_higher_education, data=data)
a18 = lm(final_grade ~ failures*weekday_alcohol_cons, data=data)
a19 = lm(final_grade ~ extra_school_support*wants_higher_education, data=data)
a20 = lm(final_grade ~ extra_school_support*weekday_alcohol_cons, data=data)
a21 = lm(final_grade ~ wants_higher_education*weekday_alcohol_cons, data=data)
anova(a1, a2, a3, a4, a5, a6, a7, a8, a9, a10, a11, a12, a13, a15, a16, a17, a18, a19, a20, a21)



b01 = lm(final_grade~school*failures + mom_edu*wants_higher_education, data=data)
b02 = lm(final_grade~school*failures + mom_edu*extra_school_support, data=data)
b03 = lm(final_grade~school*failures + mom_edu*weekday_alcohol_cons, data=data)
b04 = lm(final_grade~school*failures + extra_school_support*wants_higher_education, data=data)
b05 = lm(final_grade~school*failures + extra_school_support*weekday_alcohol_cons, data=data)
b06 = lm(final_grade~school*failures + wants_higher_education*weekday_alcohol_cons, data=data)
anova(b01, b02, b03, b04, b05, b06)
#Nope Everything is done










data %>% group_by(failures, mom_edu, school) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=failures, x=mom_edu, color=school, size = mean_final_grade), alpha=0.5)
data %>% group_by(failures, mom_edu, school) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=extra_school_support, x=wants_higher_education, color=school, size = mean_final_grade), alpha=0.5)
data %>% group_by(extra_school_support, wants_higher_education, school) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=extra_school_support, x=wants_higher_education, color=school, size = mean_final_grade), alpha=0.5)
data %>% group_by(failures, mom_edu) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=extra_school_support, x=wants_higher_education, size = mean_final_grade), alpha=0.5)
data %>% group_by(failures, mom_edu) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=failures, x=mom_edu, size = mean_final_grade), alpha=0.5)
data %>% group_by(school, mom_edu) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=school, x=mom_edu, size = mean_final_grade), alpha=0.5)
data %>% group_by(extra_school_support, mom_edu) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=extra_school_support, x=mom_edu, size = mean_final_grade), alpha=0.5)
data %>% group_by(wants_higher_education, mom_edu) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=wants_higher_education, x=mom_edu, size = mean_final_grade), alpha=0.5)
data %>% group_by(school, failures) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=school, x=failures, size = mean_final_grade), alpha=0.5)
data %>% group_by(extra_school_support, failures) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=extra_school_support, x=failures, size = mean_final_grade), alpha=0.5)
data %>% group_by(extra_school_support, wants_higher_education) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=extra_school_support, x=wants_higher_education, size = mean_final_grade), alpha=0.5)
data %>% group_by(failures, wants_higher_education) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=failures, x=wants_higher_education, size = mean_final_grade), alpha=0.5)
data %>% group_by(mom_edu, extra_school_support, wants_higher_education) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=extra_school_support, x=mom_edu, color=wants_higher_education, size = mean_final_grade), alpha=0.5)
data %>% group_by(mom_edu, extra_school_support, wants_higher_education) %>% summarise(mean_final_grade = mean(final_grade)) %>% ggplot() + geom_point(aes(y=wants_higher_education, x=mom_edu, color=extra_school_support, size = mean_final_grade), alpha=0.5)
#mean grades by variables




data2 = data %>% mutate(mom_edu=factor(mom_edu), dad_edu=factor(dad_edu), traveltime=factor(traveltime), studytime=factor(studytime), failures=factor(failures), family_relationship=factor(family_relationship), freetime=factor(freetime), going_out_amount=factor(going_out_amount), weekday_alcohol_cons=factor(weekday_alcohol_cons), weekend_alcohol_cons=factor(weekend_alcohol_cons), health=factor(health))
ffit_res <- as.tbl(data.frame(fitted=fitted(fstep), res=residuals(fstep)))
data_ffit_res <- as.tbl(cbind(data, ffit_res))
#residuals from data as factors


cschool = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=school)) + geom_point(alpha=0.5)
csex = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=sex)) + geom_point(alpha=0.5)
cage = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=age)) + geom_point(alpha=0.5)
curban_rural = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=urban_rural)) + geom_point(alpha=0.5)
cfam_size = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=fam_size)) + geom_point(alpha=0.5)
cparents_cohabit = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=parents_cohabit)) + geom_point(alpha=0.5)
cmom_edu = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=mom_edu)) + geom_point(alpha=0.5)
cdad_edu = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=dad_edu)) + geom_point(alpha=0.5)
cmom_job = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=mom_job)) + geom_point(alpha=0.5)
cdad_job = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=dad_job)) + geom_point(alpha=0.5)
cschool_reason = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=school_reason)) + geom_point(alpha=0.5)
cstudent_guardian = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=student_guardian)) + geom_point(alpha=0.5)
ctraveltime = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=traveltime)) + geom_point(alpha=0.5)
cstudytime = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=studytime)) + geom_point(alpha=0.5)
cfailures = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=failures)) + geom_point(alpha=0.5)
cextra_school_support = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=extra_school_support)) + geom_point(alpha=0.5)
cfam_edu_support = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=fam_edu_support)) + geom_point(alpha=0.5)
cpaid_extra_classes = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=paid_extra_classes)) + geom_point(alpha=0.5)
cextracurricular_activities = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=extracurricular_activities)) + geom_point(alpha=0.5)
cattended_nursery = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=attended_nursery)) + geom_point(alpha=0.5)
cwants_higher_education = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=wants_higher_education)) + geom_point(alpha=0.5)
cinternet = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=internet)) + geom_point(alpha=0.5)
chas_romantic_partner = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=has_romantic_partner)) + geom_point(alpha=0.5)
cfamily_relationship = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=family_relationship)) + geom_point(alpha=0.5)
cfreetime = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=freetime)) + geom_point(alpha=0.5)
cgoing_out_amount = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=going_out_amount)) + geom_point(alpha=0.5)
cweekday_alcohol_cons = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=weekday_alcohol_cons)) + geom_point(alpha=0.5)
cweekend_alcohol_cons = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=weekend_alcohol_cons)) + geom_point(alpha=0.5)
chealth = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=health)) + geom_point(alpha=0.5)
cabsences = data_ffit_res %>% ggplot(aes(x=fitted, y=res, color=absences)) + geom_point(alpha=0.5)

grid.arrange(cschool, csex, cage, curban_rural, cfam_size, cparents_cohabit, cmom_edu, cdad_edu, cmom_job, cdad_job, cschool_reason, cstudent_guardian, ctraveltime, cstudytime, cfailures, cextra_school_support, cfam_edu_support, cpaid_extra_classes, cextracurricular_activities, cattended_nursery, cwants_higher_education, cinternet, chas_romantic_partner, cfamily_relationship, cfreetime, cgoing_out_amount, cweekday_alcohol_cons, cweekend_alcohol_cons, chealth, cabsences, nrow)
#Residuals with different things as the color var
#c for colors

















```



###Grade vs Variable with Alpha & Jitter

```{r}
pschool = data2 %>% ggplot(aes(x=school, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
psex = data2 %>% ggplot(aes(x=sex, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
page = data2 %>% ggplot(aes(x=age, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
purban_rural = data2 %>% ggplot(aes(x=urban_rural, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pfam_size = data2 %>% ggplot(aes(x=fam_size, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pparents_cohabit = data2 %>% ggplot(aes(x=parents_cohabit, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pmom_edu = data2 %>% ggplot(aes(x=mom_edu, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pdad_edu = data2 %>% ggplot(aes(x=dad_edu, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pmom_job = data2 %>% ggplot(aes(x=mom_job, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pdad_job = data2 %>% ggplot(aes(x=dad_job, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pschool_reason = data2 %>% ggplot(aes(x=school_reason, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pstudent_guardian = data2 %>% ggplot(aes(x=student_guardian, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
ptraveltime = data2 %>% ggplot(aes(x=traveltime, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pstudytime = data2 %>% ggplot(aes(x=studytime, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pfailures = data2 %>% ggplot(aes(x=failures, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pextra_school_support = data2 %>% ggplot(aes(x=extra_school_support, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pfam_edu_support = data2 %>% ggplot(aes(x=fam_edu_support, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
ppaid_extra_classes = data2 %>% ggplot(aes(x=paid_extra_classes, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pextracurricular_activities = data2 %>% ggplot(aes(x=extracurricular_activities, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pattended_nursery = data2 %>% ggplot(aes(x=attended_nursery, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pwants_higher_education = data2 %>% ggplot(aes(x=wants_higher_education, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pinternet = data2 %>% ggplot(aes(x=internet, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
phas_romantic_partner = data2 %>% ggplot(aes(x=has_romantic_partner, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pfamily_relationship = data2 %>% ggplot(aes(x=family_relationship, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pfreetime = data2 %>% ggplot(aes(x=freetime, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pgoing_out_amount = data2 %>% ggplot(aes(x=going_out_amount, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pweekday_alcohol_cons = data2 %>% ggplot(aes(x=weekday_alcohol_cons, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pweekend_alcohol_cons = data2 %>% ggplot(aes(x=weekend_alcohol_cons, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
phealth = data2 %>% ggplot(aes(x=health, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)
pabsences = data2 %>% ggplot(aes(x=absences, y=final_grade)) + geom_point(alpha=0.1) + geom_jitter(position = position_jitter(width = .1, height=0), alpha=0.06)


grid.arrange(pschool, psex, page, purban_rural, pfam_size, pparents_cohabit, pmom_edu, pdad_edu, pmom_job, pdad_job, pschool_reason, pstudent_guardian, ptraveltime, pstudytime, pfailures, pextra_school_support, pfam_edu_support, ppaid_extra_classes, pextracurricular_activities, pattended_nursery, pwants_higher_education, pinternet, phas_romantic_partner, pfamily_relationship, pfreetime, pgoing_out_amount, pweekday_alcohol_cons, pweekend_alcohol_cons, phealth, pabsences)
#final grade vs each variable, with some horizontal jitter
#p for points
```


```{r}
ggplot(data_ffit_res, aes(y = final_grade, x = fitted)) + geom_point(alpha=0.2) + geom_abline(intercept=0, slope=1, colour="red")
#fitted vs final grade
#fitted data vs final grade, y=x line











vschool = data2 %>% ggplot(aes(x=school, y=final_grade)) + geom_violin()
vsex = data2 %>% ggplot(aes(x=sex, y=final_grade)) + geom_violin()
vage = data2 %>% ggplot(aes(x=age, y=final_grade)) + geom_violin()
vurban_rural = data2 %>% ggplot(aes(x=urban_rural, y=final_grade)) + geom_violin()
vfam_size = data2 %>% ggplot(aes(x=fam_size, y=final_grade)) + geom_violin()
vparents_cohabit = data2 %>% ggplot(aes(x=parents_cohabit, y=final_grade)) + geom_violin()
vmom_edu = data2 %>% ggplot(aes(x=mom_edu, y=final_grade)) + geom_violin()
vdad_edu = data2 %>% ggplot(aes(x=dad_edu, y=final_grade)) + geom_violin()
vmom_job = data2 %>% ggplot(aes(x=mom_job, y=final_grade)) + geom_violin()
vdad_job = data2 %>% ggplot(aes(x=dad_job, y=final_grade)) + geom_violin()
vschool_reason = data2 %>% ggplot(aes(x=school_reason, y=final_grade)) + geom_violin()
vstudent_guardian = data2 %>% ggplot(aes(x=student_guardian, y=final_grade)) + geom_violin()
vtraveltime = data2 %>% ggplot(aes(x=traveltime, y=final_grade)) + geom_violin()
vstudytime = data2 %>% ggplot(aes(x=studytime, y=final_grade)) + geom_violin()
vfailures = data2 %>% ggplot(aes(x=failures, y=final_grade)) + geom_violin()
vextra_school_support = data2 %>% ggplot(aes(x=extra_school_support, y=final_grade)) + geom_violin()
vfam_edu_support = data2 %>% ggplot(aes(x=fam_edu_support, y=final_grade)) + geom_violin()
vpaid_extra_classes = data2 %>% ggplot(aes(x=paid_extra_classes, y=final_grade)) + geom_violin()
vextracurricular_activities = data2 %>% ggplot(aes(x=extracurricular_activities, y=final_grade)) + geom_violin()
vattended_nursery = data2 %>% ggplot(aes(x=attended_nursery, y=final_grade)) + geom_violin()
vwants_higher_education = data2 %>% ggplot(aes(x=wants_higher_education, y=final_grade)) + geom_violin()
vinternet = data2 %>% ggplot(aes(x=internet, y=final_grade)) + geom_violin()
vhas_romantic_partner = data2 %>% ggplot(aes(x=has_romantic_partner, y=final_grade)) + geom_violin()
vfamily_relationship = data2 %>% ggplot(aes(x=family_relationship, y=final_grade)) + geom_violin()
vfreetime = data2 %>% ggplot(aes(x=freetime, y=final_grade)) + geom_violin()
vgoing_out_amount = data2 %>% ggplot(aes(x=going_out_amount, y=final_grade)) + geom_violin()
vweekday_alcohol_cons = data2 %>% ggplot(aes(x=weekday_alcohol_cons, y=final_grade)) + geom_violin()
vweekend_alcohol_cons = data2 %>% ggplot(aes(x=weekend_alcohol_cons, y=final_grade)) + geom_violin()
vhealth = data2 %>% ggplot(aes(x=health, y=final_grade)) + geom_violin()
vabsences = data2 %>% ggplot(aes(x=absences, y=final_grade)) + geom_violin()

grid.arrange(vschool, vsex, vage, vurban_rural, vfam_size, vparents_cohabit, vmom_edu, vdad_edu, vmom_job, vdad_job, vschool_reason, vstudent_guardian, vtraveltime, vstudytime, vfailures, vextra_school_support, vfam_edu_support, vpaid_extra_classes, vextracurricular_activities, vattended_nursery, vwants_higher_education, vinternet, vhas_romantic_partner, vfamily_relationship, vfreetime, vgoing_out_amount, vweekday_alcohol_cons, vweekend_alcohol_cons, vhealth, vabsences)

#violin plot of final grade vs all the variables
#v for violin










dschool = data %>% ggplot(aes(x=school)) + geom_histogram()
dsex = data %>% ggplot(aes(x=sex)) + geom_histogram()
dage = data %>% ggplot(aes(x=age)) + geom_histogram()
durban_rural = data %>% ggplot(aes(x=urban_rural)) + geom_histogram()
dfam_size = data %>% ggplot(aes(x=fam_size)) + geom_histogram()
dparents_cohabit = data %>% ggplot(aes(x=parents_cohabit)) + geom_histogram()
dmom_edu = data %>% ggplot(aes(x=mom_edu)) + geom_histogram()
ddad_edu = data %>% ggplot(aes(x=dad_edu)) + geom_histogram()
dmom_job = data %>% ggplot(aes(x=mom_job)) + geom_histogram()
ddad_job = data %>% ggplot(aes(x=dad_job)) + geom_histogram()
dschool_reason = data %>% ggplot(aes(x=school_reason)) + geom_histogram()
dstudent_guardian = data %>% ggplot(aes(x=student_guardian)) + geom_histogram()
dtraveltime = data %>% ggplot(aes(x=traveltime)) + geom_histogram()
dstudytime = data %>% ggplot(aes(x=studytime)) + geom_histogram()
dfailures = data %>% ggplot(aes(x=failures)) + geom_histogram()
dextra_school_support = data %>% ggplot(aes(x=extra_school_support)) + geom_histogram()
dfam_edu_support = data %>% ggplot(aes(x=fam_edu_support)) + geom_histogram()
dpaid_extra_classes = data %>% ggplot(aes(x=paid_extra_classes)) + geom_histogram()
dextracurricular_activities = data %>% ggplot(aes(x=extracurricular_activities)) + geom_histogram()
dattended_nursery = data %>% ggplot(aes(x=attended_nursery)) + geom_histogram()
dwants_higher_education = data %>% ggplot(aes(x=wants_higher_education)) + geom_histogram()
dinternet = data %>% ggplot(aes(x=internet)) + geom_histogram()
dhas_romantic_partner = data %>% ggplot(aes(x=has_romantic_partner)) + geom_histogram()
dfamily_relationship = data %>% ggplot(aes(x=family_relationship)) + geom_histogram()
dfreetime = data %>% ggplot(aes(x=freetime)) + geom_histogram()
dgoing_out_amount = data %>% ggplot(aes(x=going_out_amount)) + geom_histogram()
dweekday_alcohol_cons = data %>% ggplot(aes(x=weekday_alcohol_cons)) + geom_histogram()
dweekend_alcohol_cons = data %>% ggplot(aes(x=weekend_alcohol_cons)) + geom_histogram()
dhealth = data %>% ggplot(aes(x=health)) + geom_histogram()
dabsences = data %>% ggplot(aes(x=absences)) + geom_histogram()

grid.arrange(dschool, dsex, dage, durban_rural, dfam_size, dparents_cohabit, dmom_edu, ddad_edu, dmom_job, ddad_job, dschool_reason, dstudent_guardian, dtraveltime, dstudytime, dfailures, dextra_school_support, dfam_edu_support, dpaid_extra_classes, dextracurricular_activities, dattended_nursery, dwants_higher_education, dinternet, dhas_romantic_partner, dfamily_relationship, dfreetime, dgoing_out_amount, dweekday_alcohol_cons, dweekend_alcohol_cons, dhealth, dabsences)
#d for 'density' but actually histograms
#The only thing that's log-ish are absences and family relations, neither of which play a big role so log transforms are whatever

```
