---
title: "Final_Markdown"
author: "Mihai Sirbu"
date: "Monday, May 11, 2015"
output: pdf_document
---

```{r, echo=FALSE, include=FALSE}
library(dplyr)
```

###Read in Data

```{r}
data = as.tbl(read.csv("student-por.csv", sep=";", stringsAsFactors=TRUE))
```

###Rename the Variables

```{r}

rename_cols <- c(4:12,16:21,23:24,26:28,31:33)

colnames(data)[rename_cols] = c("urban_rural",
                                "fam_size",
                                "parents_cohabit",
                                "mom_edu",
                                "dad_edu",
                                "mom_job",
                                "dad_job",
                                "school_reason",
                                "student_guardian",
                                "extra_school_support",
                                "fam_edu_support",
                                "paid_extra_classes",
                                "extracurricular_activities",
                                "attended_nursery",
                                "wants_higher_education",
                                "has_romantic_partner",
                                "family_relationship",
                                "going_out_amount",
                                "weekday_alcohol_cons",
                                "weekend_alcohol_cons",
                                "grade_one",
                                "grade_two",
                                "final_grade")
```

###Meaningful Variable Values

```{r}
data <-  data %>% 
  mutate(school = ifelse(school=='GP', 'Gabriel Pereira', 'Mousinho da Silveira')) %>%
  mutate(urban_rural = ifelse(urban_rural=='R', 'Rural', 'Urban')) %>% 
  mutate(fam_size = ifelse(fam_size=='GT3', '>3', '<=3')) %>%
  mutate(parents_cohabit = ifelse(parents_cohabit=='A', 'Apart', 'Together')) %>%
  mutate_each_(funs(ifelse(.=='yes', TRUE, FALSE)), 
               c("extra_school_support", "fam_edu_support", "paid_extra_classes", 
                 "extracurricular_activities", "attended_nursery", 
                 "wants_higher_education", "internet", "has_romantic_partner"))
  
```


Summary: 1) Describe the task you performed (in one sentence). 2) Describe the data you used (in one sentence). 3) Describe the analysis method you used (in one sentence). 4) Summarize your results (in one or two sentences).

(1) The goal of this project was to investigate and understand the relationship between student demographics and academic achievemet (based on final grade) in a language class in Portugal. 
(2) 
(3) To investigate this relationship we decided to use linear regression and we believed this would allow us to interpret our results more easily. 
(4)


Introduction (1/2 page): Describe the task you performed in more detail. Include motivation for performing this task.

Materials and Methods (1.5-2 pages): Describe the dataset you used in more detail (you can get this from Step 3) Describe analysis methods you used, and how you are evaluating performance (from Step 3)

Results (about 3 pages): Describe the results you found. Make sure you quantify your findings (if inference, provide interpretation of resulting model. E.g., we found that house prices increase by $10,000 (P<0.05) on average when they are within 5 miles of a university campus; if prediction, we found we could predict house prices with mean squared error of $10,000.

Conclusion (1/2 page): Concluding remarks summarizing findings and discussing possible improvements to your analysis.

